# Well Data -- Data Loggers

**Updated: `r format(Sys.time(), '%Y %B %d')`**

## Introduction

This section provides basic QA/QC assessments for the well logger data.

Analyses aim to characterize basic data structure and identify potential issues, clean data as needed, then produce an archive-worthy output data.

Because they originate as separate source files, 2016-2018 data are brought in and examined separately. In addition, data are cleaned to produce archival files for further analyses.

```{r,echo=FALSE}
# library(here)
# here()
# install.packages("bindrcpp")
suppressPackageStartupMessages(library(tidyverse))
library(fs)
suppressPackageStartupMessages(library(sf))
# library(raster)
suppressPackageStartupMessages(library(janitor))
suppressPackageStartupMessages(library(readxl))
# library(glue)
suppressPackageStartupMessages(library(mapview))
# library(ggmap)
# library(ggrepel)
suppressPackageStartupMessages(library(viridis))
library(ggExtra)
library(DT)
library(kableExtra)
suppressPackageStartupMessages(library(lubridate))
library(anytime)
suppressPackageStartupMessages(library(compare))
suppressPackageStartupMessages(library(skimr)) ## some useful functions
suppressPackageStartupMessages(library(dataMaid))

```


### 2012-2015 (DCC Collection)

#### Data import and initial inspection 
```{r}
# import of tsv 
log.dat.raw <- readr::read_tsv("data/NSF_DataArchive20180208/Well_Logger_Data_2012-2015/Yell_Well_Data_Logger2012_2015.txt")

log.dat <- log.dat.raw %>% 
  clean_names() %>% 
  tibble::rownames_to_column(var = "RecordID") 

```

```{r}
## evaluate

log.dat %>% glimpse()


```


> **Questions/Issues:**    
> 1. There was some bad exporting in the DCC file. No useable time stamps -- everything is characterer in date format.     
> 2. May wish to find the orginal logger data. This may work for daily aggregated data. Otherwise, not sure what to do with the 6 per-day records.  

```{r}
## data type changes
log.dat <- log.dat %>% 
  mutate(date = mdy(date))


## what are the distinct wells?
log.dat %>% 
  distinct(site_logger)

# note there are 19 distinct "site_logger" values and 19 distinct "serial_num"
log.dat %>% 
  distinct(serial_num)

```

```{r}
dly.mean <- log.dat %>% 
  group_by(site,site_logger, wat_id2, date) %>% 
  summarise(mean.dly.abslevel = mean(abslevel,na.rm = TRUE), mean.water_column_m = mean(water_column_m, na.rm = TRUE))

## add in year and mob=nth columns
dly.mean <- dly.mean %>% 
  mutate(yr = lubridate::year(date)) %>% 
  mutate(month = lubridate::month(date)) %>% 
  mutate(doy = lubridate::yday(date)) %>% 
  mutate(doy.d = format(date, format="%m-%d"))


dly.mean %>% 
  ggplot(aes(x=date,y=mean.water_column_m)) +
  geom_line(aes(color=site_logger)) +
  labs(title = "Mean daily 'water_column_m' values") +
  facet_wrap(~site)

dly.mean %>% 
  ggplot(aes(x=date,y=mean.dly.abslevel)) +
  geom_line(color="blue") +
  labs(title = "Mean daily 'abslevel' values") +
  facet_wrap(~site, scales = "free_y")

  
```

```{r, fig.height=15}
## plot daily traces

dly.mean %>% # names()
  filter(month > 2 & month < 10) %>% 
  mutate(yr = as.factor(yr)) %>% 
  # ggplot(aes(x=doy,y=mean.water_column_m)) +
  ggplot(aes(x=doy,y=mean.dly.abslevel)) +
  geom_line(aes(color=yr)) +
  labs(title = "Mean daily 'water_column_m' values") +
  facet_wrap(~site_logger, scale = 'free_y', ncol = 3) +
  theme_minimal()


```



### 2016 Data 
#### Data import and initial inspection 

```{r}

```




### 2017 Data 
#### Data import and initial inspection 

```{r}

```


### 2018 Data 
#### Data import and initial inspection 

```{r}

```


### Clean and Combine Years

### Export Archivable Version of Cleaned Data for Further Analysis