---
output: 
  html_document: 
    theme: readable
---
# Environmental Logger Data 2012 - 2019

**Updated: `r format(Sys.time(), '%Y %B %d')`**


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

```

## Functions
```{r}
## functions
### quick ref

#- names to gt

names2gt <- function(df){
  names(df) %>% 
    enframe() %>% 
    gt() %>% 
    tab_header(title = "field names")
}

### plotting

#--doy vs vwc
pl_temp_exp_doy <- function(df){
  df %>% 
  filter(temp_c > 0) %>% 
  filter()
  mutate(yr = as.factor(yr)) %>% 
  ggplot(aes(doy,temp_c)) +
  geom_line(aes(color = yr)) +
  facet_grid(plot~site) +
  theme_minimal() +
  labs(x = "Day of year", y = "Volumetric water content")
}

#--doy vs vwc
pl_vwc_doy1 <- function(df){
  df %>% 
  mutate(yr = as.factor(yr)) %>% 
  ggplot(aes(doy,vwc)) +
  geom_line(aes(color = yr)) +
  facet_grid(plot~site) +
  theme_minimal() +
  labs(x = "Day of year", y = "Volumetric water content")
}

pl_vwc_doy2 <- function(df){
  df %>% 
  mutate(yr = as.factor(yr)) %>% 
  ggplot(aes(doy,vwc)) +
  geom_line(aes(color = yr)) +
  facet_grid(site~plot) +
  theme_minimal() +
  labs(x = "Day of year", y = "Volumetric water content")
}


pl_vwc_date <- function(df){
  df %>% 
  mutate(yr = as.factor(yr)) %>% 
  ggplot(aes(measurement_time,vwc)) +
  geom_line(aes(color = var_vwc)) +
  facet_wrap(~year, scales = "free_x") +
  theme_minimal() +
  labs(x = "Day of year", y = "Volumetric water content")
}

#--annual time series


### summary tables


```

## Introduction

This section provides a basic interrogation of environmental measurement data archived on the "Digital Collections of Colorado" (DCC) library site and data collected in 2016-2018. These analyses are aimed at characterizing basic data structure and identifying potential issues such as:

* Inconsistently named factors such as site name or logger ID    
* Missing values  
* Data values outside of expected range or showing unusual patterns


```{r,echo=FALSE}

# library(here)
# here()
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(fs))
suppressPackageStartupMessages(library(sf))
suppressPackageStartupMessages(library(janitor))
suppressPackageStartupMessages(library(readxl))
suppressPackageStartupMessages(library(mapview))
# library(ggmap)
# library(ggrepel)
suppressPackageStartupMessages(library(viridis))
library(ggExtra)
library(DT)
suppressPackageStartupMessages(library(lubridate))
library(anytime)
# suppressPackageStartupMessages(library(compare))
suppressPackageStartupMessages(library(skimr)) ## some useful functions
suppressPackageStartupMessages(library(dataMaid))
library(visdat)
library(gt)
library(tsibble) ## ts
library(feasts) ## ts

# install.packages("devtools")
# devtools::install_github("twitter/AnomalyDetection")

library(AnomalyDetection)

```

```{r}

```


## Soil Moisture data 

```{r}
### 2012-2015 Data (DCC Collection)

### 2012-2015 (DCC Collection)

#### Data import and initial inspection 

### 2016 Data 

#### Data import and initial inspection 
```

```{r, eval=TRUE}

# use 'readxls' functions
# read_excel(path, sheet = NULL, range = NULL, col_names = TRUE,
#   col_types = NULL, na = "", trim_ws = TRUE, skip = 0, n_max = Inf,
#   guess_max = min(1000, n_max))

#  Get the names and paths. 
list.files <- fs::dir_ls("data/raw/env_data_loggers",recurse = TRUE, glob = "*.xls")

## Purrr: take list and read in first sheet of each folder
env1 <- list.files %>%
  # map_df(read_excel, .id = "file", sheet = 1,skip = 2, col_types = c("text","date", "numeric", "numeric","numeric","numeric","numeric","numeric")) %>%
  map_df(read_excel, .id = "file", sheet = 1,skip = 2, col_types = "text") %>% # have to specify the "text" otherwise it fubars the guessing of data types with the date
  janitor::clean_names() 

```

```{r}
env1 %>% 
  distinct(file) %>% 
  datatable(caption = "Distinct decagon logger files to merge")
```

**158 distinct logger files to deal with...**

```{r, eval=TRUE, echo=TRUE}

####
env1 <- env1 %>% 
  mutate(date = excel_numeric_to_date(as.numeric(as.character(measurement_time)), date_system = "modern"))  %>%
  mutate(yr = as.integer(year(date))) %>% 
  mutate_at(vars(contains("temp")), as.numeric) %>% 
  mutate_at(vars(contains("vwc")), as.numeric) %>% 
  mutate(mm_precip = as.numeric(mm_precip)) %>% 
  mutate_if(is.character, tolower) %>% 
  select(-contains("none"))

env1 %>% 
  sample_n(10000) %>% 
  vis_dat()

env1 %>% 
  distinct(file) %>% datatable()

```

```{r}
# split 
env1 <- env1 %>% 
  separate(col = file, into = c("a1","a2","a3","a4","a5"), sep = "/",fill = "right", remove = FALSE) %>% 
  select(-c(a1,a2,a3,a4)) %>%
  # str_split(fruits, " and ")
  mutate(file_n = str_remove(string = a5,pattern = ".xls")) 

# filter out the 'archived' files
env1 <- env1 %>%
  filter(!str_detect(file_n, "Arch")) %>% #filter out the "archived" files
  filter(!str_detect(file_n, "arch"))

env1 <- env1 %>% #filter out the "archived" files
  mutate(file_n = str_replace(string = file_n,pattern = "-",replacement = " ")) %>% 
  mutate(file_n = str_replace(string = file_n,pattern = "-",replacement = " ")) %>% 
  # View()
  # distinct(file_n) %>% datatable()
  separate(col = file_n, into = c("site","datep","logger_info"),sep = " ",remove = FALSE) 

## to lower
env1 <- env1 %>%
  mutate_if(.predicate = is.character,.funs = tolower)

```

```{r}
env1 %>% 
  # View()
  tabyl(logger_info,site) %>% 
  gt() %>% 
  tab_header(title = "distinct file suffix and site")
```

```{r}
env1 <- env1 %>% 
  select(-contains('none')) %>% 
  select(-c(a5))

env1 %>% 
  names2gt()

```


# Lookup join
```{r}
## join in site plot info

# env1 %>% 
#   distinct(site) %>% 
#   write_csv("./data/lu_em50_site_plot.csv")

lu.em50 <- read_csv("./data/lu_em50_site_plot.csv")

##
env1 <- env1 %>% 
  left_join(.,lu.em50,by='site') 

# env1 %>%
#   # distinct(site2)
#   filter(is.na(site2)) %>% 
#   distinct(file_n)

## to addres the missing site2 attributes...
env1 <- env1 %>%
  mutate(site2 = case_when(file_n == "eb2obs 24oct2018 0841" ~ "eb2",
                           file_n == "eb2obs 08jan2020 1758" ~ "eb2",
                           TRUE ~ site2))

env1 <- env1 %>% 
  mutate(site = site2) %>% 
  select(-site2)

env1 <- env1 %>% 
  mutate(plot = case_when(is.na(plot) ~ "obs",
                          TRUE ~ plot))


env1 <- env1 %>% 
  rename(logger_id = logger_info)
```

```{r}

env1 %>%
  # distinct(site)
  # filter(is.na(site)) %>% 
  # distinct(file) %>%  datatable()
  tabyl(site,plot) %>%
  gt() 

```

no eb1 dc!

```{r}
env1 %>% 
  sample_n(10000) %>% 
  vis_dat()

```

```{r}
env1 %>% 
  tabyl(site, yr) %>% datatable(caption = "issues with date parsing at 'Rose'")

env1 %>% 
  tabyl(logger_id, yr) %>% datatable(caption = "issues with date parsing at 'Rose'")
```


```{r}

#### excel time fix
env1 <- env1 %>% 
  mutate(measurement_time = as.numeric(measurement_time)) %>%
  mutate(measurement_time = janitor::excel_numeric_to_date(measurement_time, date_system = "modern", include_time = TRUE))

env1 <- env1 %>% 
  mutate(doy = lubridate::yday(measurement_time)) %>% 
  mutate(month = lubridate::month(measurement_time)) %>% 
  mutate(hour = lubridate::hour(measurement_time)) 



```


# rose 2050 and other issues with problem time stamps

```{r}
# probelm dates
env1 %>% 
  filter(!(yr>2001 & yr <2050)) %>% 
  mutate(site_plot = paste0(site,"-",plot)) %>% 
  distinct(file,site) %>%
  # write_csv("env_sensor_date_issues_2qc.csv")
  # filter(site == "rose") %>% 
  datatable()

```


```{r}

## clean
env1 <- env1 %>% 
  filter(yr>2001 & yr <2050) %>% 
  mutate(site_plot = paste0(site,"-",plot))
```


```{r, eval=FALSE}
################# trying on a subset...
## just the august 17 folder. Trying to see if this helps...
list.files <- fs::dir_ls("data/raw/env_data_loggers/Aug 2017",recursive = TRUE, glob = "*.xls")

## Purrr: take list and read in first sheet of each folder
env1 <- list.files %>%
  # map_df(read_excel, .id = "file", sheet = 1,skip = 2, col_types = c("text","date", "numeric", "numeric","numeric","numeric","numeric","numeric")) %>%
  map_df(read_excel, .id = "file", sheet = 2, skip = 2) %>%
  # map_df(read_excel, .id = "file", sheet = 1,skip = 2, col_types = "text") %>% # have to specify the "text" otherwise it fubars the guessing of data types with the date
  janitor::clean_names() 

env1 <- env1 %>% 
  mutate_at(.vars = c(3:11),.funs = as.numeric) 



env1 %>% 
  ggplot(aes(measurement_time,m³m³vwc)) +
  geom_line() +
  facet_wrap(~file)

#env1 <- env1 %>%
  mutate(measurement_time = as.numeric(measurement_time)) %>%
  mutate(measurement_time = janitor::excel_numeric_to_date(measurement_time, date_system = "modern", include_time = TRUE))

```

```{r}
# install.packages("tsibble")
# library(tsibble)

```

## tidy temp

```{r}
## tidy

env1.temp.tidy <- env1 %>% 
  pivot_longer(
  cols = contains("temp"),
  # cols = starts_with("temp"),
  names_to = "var_temp",
  values_to = "temp_c",
  values_drop_na = TRUE
 ) 

env1.ppt.tidy <- env1.temp.tidy %>% 
  select(-contains("vwc")) 


env1.temp.tidy <- env1.temp.tidy %>% 
  select(-contains("vwc")) %>% 
  select(-c(datep, mm_precip)) 
  
env1.temp.tidy %>% 
  names()

env1.temp.tidy <- env1.temp.tidy %>%
  # select(-file) %>% 
  distinct()

env1.temp.tidy <- env1.temp.tidy %>% 
  filter(!is.na(temp_c)) %>% 
  mutate(site_logger_id = paste0(site,"-",logger_id))

```

```{r, tidytempclean}

env1.temp.tidy.clean <- env1.temp.tidy %>% 
  filter(yr >2001)

env1.temp.tidy.clean %>% 
  sample_n(10000) %>% 
  visdat::vis_dat()

```

```{r}
env1.temp.tidy.clean %>% 
  glimpse()
  
```



## temp plots

```{r}
env1.temp.tidy %>% 
  tabyl(site_logger_id,var_temp) %>% 
  gt()
```


## April through September
```{r, nest_temp}

env1.temp.tidy.clean.dly <- env1.temp.tidy.clean %>% 
  filter(temp_c > 0) %>% 
  group_by(date, month, plot, site, site_plot, yr, var_temp) %>%
  summarise(temp_c = median(temp_c, na.rm = TRUE))  
  
### nest
temp.nest.site <- env1.temp.tidy.clean.dly %>%
  filter(month > 4 & month < 11) %>% 
  group_by(site_plot) %>% 
  nest()

### functions
ggTempSite <- function(data){
  data %>% 
  ggplot(aes(date,temp_c)) +
  geom_line(aes(color = var_temp)) +
  geom_point(aes(color = var_temp)) +
  # geom_point() +
  facet_wrap(~yr, scale = "free_x", ncol = 2) +
  theme_minimal() +
  labs(x = "Date", x = "Temperature (deg C)")
}

ggTempSite_ncol1 <- function(data){
  pt1 <- data %>% 
  ggplot(aes(date,temp_c)) +
  geom_line(aes(color = var_temp)) +
  geom_point(aes(color = var_temp)) +
  # geom_point() +
  facet_wrap(~yr, scale = "free_x", ncol = 1) +
  theme_minimal() +
  labs(x = "Date", y = "Temperature (deg C)")
}

```

```{r}
temp.nest.site <- temp.nest.site %>%
  mutate(ggp_site = map(.x = data, .f = ggTempSite)) %>% 
  mutate(ggp_siteb = map(.x = data, .f = ggTempSite_ncol1)) 

temp.nest.site %>% 
  pluck(4) %>% 
  pluck(2)
  
# map2(paste0(plots$country, ".pdf"), plots$plot, ggsave)


```

```{r, eval = FALSE}
## write plots
map2(paste0(temp.nest.site$site_plot,"_soiltemp",".png"), temp.nest.site$ggp_site, ggsave)

map2(paste0(temp.nest.site$site_plot,"_soiltemp_alt",".png"), temp.nest.site$ggp_siteb, ggsave)

```


# >>>v
```{r}
## reclass the temp var
env1.temp.tidy %>% 
  mutate(var_temp2 = case_when(var_temp == 'c_temp_3' ~ "temp_d1",
                               var_temp == 'c_temp_4' ~ "temp_d2",
                               var_temp == 'c_temp_5' ~ "temp_d1",
                               TRUE ~ var_temp))
  #                              c_temp_5	c_temp_6	c_temp_7	c_temp_8	temp_c_3	temp_c_4	temp_c_5	temp_c_6))
  # mutate(var_vwc = case_when(var_vwc == "m³m³vwc_2" ~ "vwc_d1",
  #                             var_vwc == "m³m³vwc_4" ~ "vwc_d2",
  #                             var_vwc == "m³m³vwc_3" ~ "vwc_d1",
  #                             var_vwc == "m³m³vwc_5" ~ "vwc_d2",
  #                             var_vwc == "m³m³vwc_6" ~ "vwc_d1",
  #                             var_vwc == "m³m³vwc_7" ~ "vwc_d2",
  #                             TRUE ~ var_vwc)) 

```



## tidy vwc

### anomoly detection
```{r, eval=FALSE}
env1 %>%
  tabyl(site, yr) %>% 
  gt() %>% 
  tab_header(title = "Count of records by year in timestamp")

# env1 %>% distinct(site_plot)

#  https://github.com/twitter/AnomalyDetection

data(raw_data)
res = AnomalyDetectionTs(raw_data, max_anoms=0.02, direction='both', plot=TRUE)
res$plot

```


```{r}
## tidy
env1.vwc.tidy <- env1 %>% 
  pivot_longer(
  cols = contains("vwc"),
  # cols = starts_with("temp"),
  names_to = "var_vwc",
  values_to = "vwc",
  values_drop_na = TRUE
 ) 

env1.vwc.tidy <- env1.vwc.tidy %>% 
  select(-contains("temp")) %>% 
  select(-c(datep)) 

env1.vwc.tidy <- env1.vwc.tidy %>%
  # select(-file) %>% 
  # distinct() %>%
  mutate(site_logger_id = paste0(site,"-",logger_id))

## !!!
env1.vwc.tidy <- env1.vwc.tidy %>% 
  mutate(var_vwc2 = case_when(var_vwc == "m³m³vwc_2" ~ "vwc_d1",
                              var_vwc == "m³m³vwc_4" ~ "vwc_d2",
                              var_vwc == "m³m³vwc_3" ~ "vwc_d1",
                              var_vwc == "m³m³vwc_5" ~ "vwc_d2",
                              var_vwc == "m³m³vwc_6" ~ "vwc_d1",
                              var_vwc == "m³m³vwc_7" ~ "vwc_d2",
                              TRUE ~ var_vwc)) 
env1.vwc.tidy %>% 
  tabyl(var_vwc2)


```

## VWC plots

```{r}



```

```{r}
# env1.vwc.tidy  %>% 
#   tabyl(site,var_vwc) %>% 
#   gt()
```


```{r}
# 
env1.vwc.tidy %>%
  # distinct(var_vwc)
  filter(vwc < .65 & vwc > .05) %>% 
  filter(var_vwc2 == "vwc_d1") %>% 
  filter(plot != "obs") %>% 
  pl_vwc_doy1 +
  labs(title = "Sensor depth 1")

env1.vwc.tidy %>%
  # distinct(var_vwc)
  filter(vwc < .65 & vwc > .05) %>% 
  filter(var_vwc2 == "vwc_d2") %>% 
  filter(plot != "obs") %>% 
  pl_vwc_doy1 +
  labs(title = "Sensor depth 2")


```

```{r}
env1.vwc.tidy %>% 
  filter(var_vwc2 == "vwc_d2") %>%
  filter(vwc < .65 & vwc > .05) %>% 
  # sample_n(50000)
  filter(plot != "obs") %>% 
  pl_vwc_doy2


```

# for DM
```{r}
## file info
env1.vwc.tidy %>% 
  distinct(file) %>%
  write_csv("./output/env_data_4DM/list_of_decagon_files.csv")

env1.precip.tidy %>% 
  distinct(file) %>%
  gt() %>% 
  tab_header(title = "Raw environmental data logger files")

```

#### precip only
```{r}
## just the precip
env1.precip.tidy <- env1.vwc.tidy %>% 
  filter(!is.na(mm_precip))

## pull out precip
env1.precip.tidy <- env1.precip.tidy %>% 
  dplyr::select(-c(vwc, var_vwc, var_vwc2)) %>% 
  distinct()

env1.precip.tidy %>% 
  distinct()

## overview
env1.precip.tidy %>% 
  visdat::vis_dat(warn_large_data = FALSE)
# ggsave("./output/env_data_4DM/decagon_precip_overview.png", width = 6, height = 6, dpi = 200)  

##
ppt_qaqcplot <- env1.precip.tidy %>%
  ggplot(aes(x = doy, y = mm_precip)) +
  geom_line(aes(color = site_logger_id)) +
  geom_point(aes(color = site_logger_id)) +
  facet_wrap(~site_plot, scales = "free_y", ncol = 2)

env1.precip.tidy %>%
  mutate(yr = as_factor(yr)) %>% 
  ggplot(aes(x = doy, y = mm_precip)) +
  geom_line(aes(color = yr)) +
  # geom_point(aes(color = yr)) +
  facet_wrap(~site_logger_id, scales = "free_y", ncol = 2) +
  theme_minimal()
ggsave("./output/env_data_4DM/decagon_precip_line_plot4qaqc.pdf", width = 8, height = 18)

## write to csv for DM
env1.precip.tidy %>%
  write_csv("./output/env_data_4DM/decagon_precip_tidy_uncleaned.csv")


```


```{r}
vwc.tidy <- env1.vwc.tidy %>% 
  dplyr::select(-c(mm_precip))

## write to csv for DM
vwc.tidy %>%
  write_csv("./output/env_data_4DM/decagon_vwc_tidy_uncleaned.csv")


env1.vwc.tidy %>% 
  select(var_vwc,var_vwc2,vwc,file_n)
  names()
```


```{r}
env1.vwc.tidy %>%
  filter(plot != "obs") %>% 
  ggplot(aes(measurement_time,vwc)) +
  geom_line(aes(color = var_vwc)) +
  facet_wrap(~site_logger_id, ncol = 4)

```

#### soil temp
```{r, eval=FALSE}
env1.temp.tidy.clean %>% 
  write_csv("./output/env_data_4DM/soil_temp/decagon_soiltemp_uncleaned.csv")
  

env1.temp.tidy.clean <- env1.temp.tidy.clean %>%
  # names()
  group_by(logger_id, site, var_temp) %>% 
  mutate(zscore = scale(temp_c)) 


```


# ! lots of f-up dates. Here are 
```{r}
# env1.vwc.tidy %>%
#   filter(yr < 2003) %>% 
#   tabyl(file) %>% 
#   gt()

```

### clean

```{r}

env1.vwc.tidy.clean <- env1.vwc.tidy %>% 
  filter(yr > 2003) %>%
  filter(vwc < .9 & vwc > 0.05)

```


```{r, eval=FALSE}
library(anomalize) #tidy anomaly detectiom
library(tidyverse) #tidyverse packages like dplyr, ggplot, tidyr
library(coindeskr) #bitcoin price extraction from coindesk

# devtools::install_github('business-science/anomalize')

btc <- get_historic_price(start = "2017-01-01")

btc_ts <- btc %>% rownames_to_column() %>% as.tibble() %>% 
  mutate(date = as.Date(rowname)) %>% select(-one_of('rowname'))

btc_ts %>% 
  time_decompose(Price) %>%
  anomalize(remainder) %>%
  time_recompose() %>%
  plot_anomalies(time_recomposed = TRUE, ncol = 3, alpha_dots = 0.5)

btc_ts %>% 
  time_decompose(Price, method = "stl", frequency = "auto", trend = "auto") %>%
  anomalize(remainder, method = "gesd", alpha = 0.05, max_anoms = 0.2) %>%
  plot_anomaly_decomposition() 

btc_ts %>% 
  time_decompose(Price) %>%
  anomalize(remainder) %>%
  time_recompose() %>%
  filter(anomaly == 'Yes') 

env1.vwc.tidy %>% 
  # distinct(site_logger_id)
  filter(site_logger_id == "eb1-1307") %>% 
  # distinct(var_vwc)
  filter(var_vwc == "m³m³vwc_2") %>% 
  time_decompose(vwc) %>%
  anomalize(remainder) %>%
  time_recompose() %>%
  plot_anomalies(time_recomposed = TRUE, ncol = 3, alpha_dots = 0.5)



```


```{r}

env1.vwc.tidy.clean %>% 
  # glimpse()
  filter(site_logger_id == "eb1-1307") %>%
  # filter(site == "eb1") %>%
  filter(var_vwc == "m³m³vwc_2") %>% 
  ggplot(aes(measurement_time,vwc)) +
  geom_line(color = "blue", alpha = .2) +
  geom_point(color = "lightblue", alpha = .8) +
  theme_minimal() +
  facet_wrap(~yr)

env1.vwc.tidy.clean.dly <- env1.vwc.tidy.clean %>% 
  group_by(date, month, plot, site, site_plot, yr, var_vwc) %>%
  summarise(vwc = median(vwc, na.rm = TRUE))

```

```{r}
env1.vwc.tidy.clean.dly %>% 
  # filter(site_logger_id == "eb1-1307") %>% 
  # filter(var_vwc == "m³m³vwc_2") %>% 
  filter(site == "eb1") %>% 
  ggplot(aes(date,vwc)) +
  geom_line(aes(color = var_vwc)) +
  # geom_point() +
  facet_wrap(~yr, scale = "free_x", ncol = 2)

env1.vwc.tidy.clean %>%
  filter(month > 4 & month < 10) %>% 
  filter(plot != "obs") %>%
  filter(vwc <.56) %>%
  filter(yr > 2009) %>% 
  group_by(yr, site, date, doy, var_vwc) %>% 
  summarise(vwc = median(vwc)) %>% 
  ggplot(aes(doy,vwc)) +
  geom_line(aes(color = var_vwc)) +
  labs(x = "Day of year", y = "Volumetric water content") +
  # geom_point() +
  facet_grid(yr~site)
ggsave("vwc_exp_facet.png", width = 9.5, height = 6.15)

```

```{r, eval=FALSE}
library(tsibble)
as_tsibble(x = env1.vwc.tidy.clean)

```


### Nested vwc

## April through September
```{r, nest_vwca}
vwc.nest1 <- env1.vwc.tidy.clean.dly %>%
  filter(month > 4 & month < 11) %>% 
  group_by(site_plot) %>% 
  nest()

# vwc.nest1 %>% 
#   # mutate(plot = map)
#   pluck(2) %>% 
#   pluck(2)

ggVWCsite <- function(data){
  data %>% 
  ggplot(aes(date,vwc)) +
  geom_line(aes(color = var_vwc)) +
  geom_point(aes(color = var_vwc)) +
  # geom_point() +
  facet_wrap(~yr, scale = "free_x", ncol = 2) +
  theme_minimal()
}

ggVWCsite_ncol1 <- function(data){
  pt1 <- data %>% 
  ggplot(aes(date,vwc)) +
  geom_line(aes(color = var_vwc)) +
  geom_point(aes(color = var_vwc)) +
  # geom_point() +
  facet_wrap(~yr, scale = "free_x", ncol = 1) +
  theme_minimal() +
  labs(x = "Date", y = "Volumetric water content")
}


vwc.nest1 <- vwc.nest1 %>%
  mutate(ggp_site = map(.x = data, .f = ggVWCsite)) %>% 
  mutate(ggp_siteb = map(.x = data, .f = ggVWCsite_ncol1)) 


vwc.nest1 %>% 
  pluck(4) %>% 
  pluck(3)
  
# map2(paste0(plots$country, ".pdf"), plots$plot, ggsave)



```

```{r, eval=FALSE}
## save
map2(paste0(vwc.nest1$site_plot,"_alt1",".png"), vwc.nest1$ggp_site, ggsave,)
map2(paste0(vwc.nest1$site_plot,"_ncol1",".png"), vwc.nest1$ggp_siteb, ggsave)
```


```{r, eval=FALSE}

dat = data.frame(elev = round( runif(20, 100, 500), 1),
                 resp = round( runif(20, 0, 10), 1),
                 grad = round( runif(20, 0, 1), 2),
                 slp = round( runif(20, 0, 35),1),
                 lat = runif(20, 44.5, 45),
                 long = runif(20, 122.5, 123.1),
                 nt = rpois(20, lambda = 25) )
head(dat)

response = names(dat)[1:3]

expl = names(dat)[4:7]

response = set_names(response)
response

expl = set_names(expl)
expl

scatter_fun = function(x, y) {
     ggplot(dat, aes(x = .data[[x]], y = .data[[y]]) ) +
          geom_point() +
          geom_smooth(method = "loess", se = FALSE, color = "grey74") +
          theme_bw() +
          labs(x = x,
               y = y)
}

scatter_fun = function(x, y) {
     ggplot(dat, aes_string(x = x, y = y ) ) +
          geom_point() +
          geom_smooth(method = "loess", se = FALSE, color = "grey74") +
          theme_bw() 
}

scatter_fun("lat", "elev")

elev_plots = map(expl, ~scatter_fun(.x, "elev") )
elev_plots

all_plots = map(response,
                ~map(expl, scatter_fun, y = .x) )

all_plots$grad[1:2]
all_plots$grad$long
all_plots[[3]][[3]]

# pdf("all_scatterplots.pdf")
# all_plots
# dev.off()

plotnames = imap(all_plots, ~paste0(.y, "_", names(.x), ".png")) %>%
     flatten()
plotnames

walk2(plotnames, flatten(all_plots), ~ggsave(filename = .x, plot = .y, 
                                             height = 7, width = 7))
```



```{r, nest_vwc}
vwc.nest2 <- env1.vwc.tidy.clean.dly %>%
  group_by(site, plot) %>% 
  nest()

# vwc.nest1 %>% 
#   # mutate(plot = map)
#   pluck(2) %>% 
#   pluck(2)

ggVWCsite_plot <- function(data){
  data %>% 
  ggplot(aes(date,vwc)) +
  geom_line(aes(color = var_vwc)) +
  geom_point(aes(color = var_vwc)) +
  # geom_point() +
  facet_wrap(~yr, scale = "free_x", ncol = 3) +
  theme_minimal() +
  labs(x = "Date", y = "Volumetric water content")
}

vwc.nest2 <- vwc.nest2 %>%
  mutate(ggp_site = map(.x = data, .f = ggVWCsite_plot)) 

vwc.nest2 %>% 
  pluck(4) %>% 
  pluck(3)
  


```

```{r, eval = FALSE}
# map2(paste0(plots$country, ".pdf"), plots$plot, ggsave)

## save
map2(paste0(vwc.nest2$site_plot,"_nest2",".pdf"), vwc.nest1$ggp_site, ggsave)
# map2(paste0(vwc.nest1$site_plot,"_ncol1",".png"), vwc.nest1$ggp_siteb, ggsave)
```


```{r}

env1.vwc.tidy %>%
  # distinct(site)
  filter(site == "eb2") %>% 
  # filter(plot != "obs") %>% 
  ggplot(aes(measurement_time,vwc)) +
  geom_point(aes(color = var_vwc)) +
  facet_wrap(~site, ncol = 4)

wvc.median <- env1.vwc.tidy %>%
  filter(vwc < .65 & vwc > .05) %>%
  group_by(var_vwc,date, site,plot, yr) %>%
  summarise(vwc_median = median(vwc,na.rm = TRUE), vwc_max = max(vwc,na.rm = TRUE), vwc_range = IQR(vwc))


wvc.median %>% 
  filter(site == "elk") %>% 
  filter(plot != "obs") %>% 
  ggplot(aes(date,vwc_max)) +
  geom_point(aes(color = var_vwc)) +
  facet_wrap(~yr, ncol = 4, scales = "free_x")

```


```{r, eval=FALSE}
env1.vwc.tidy %>% 
  filter(plot != "obs") %>% 
  filter(yr != 2050) %>% 
  mutate(yr = as.factor(yr)) +
  ggplot(aes(doy,vwc)) +
  geom_line(aes(color = yr)) +
  facet_grid(plot~site)

env1.vwc.tidy %>% 
  filter(plot != "obs") %>% 
  filter(yr != 2050) %>% 
  mutate(yr = as.factor(yr)) +
  ggplot(aes(doy,vwc)) +
  geom_line(aes(color = yr)) +
  facet_grid(plot~site)

           
```


### Export Archivable Version of Cleaned Data for Further Analysis 


#