---
title: "YNP_logan_mtg_2020"
author: ""
date: "1/27/2020"
output: 
  powerpoint_presentation: 
    fig_height: 5.5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r}
library(tidyverse)
library(patchwork)
library(gt)
library(DT)
library(sf)

```

## Overview of data sets

### Willow Measurements
- Willow height
- Willow production and utilization

### Hydrological measurements
- Manual water level measurements
- Water level dataloggers

## Environmental data
- Soil moisture and temperature
- Site temp and precipitation

## Geospatial

### Site survey data
- Wells/staff gages
- Willows

### Ancillary GIS
- Fences
- DEM
- Beaver occupancy
- Instrument locations

## Other data
- Camera traps
- ???


## Study sites
```{r}
# maps
sites <- st_read("./data/geospatial/raw/Willow_Sites.shp", quiet = TRUE)

sites <- sites %>% 
  filter(!is.na(Type)) 

```

```{r}
mapview::mapview(sites, zcol="Type")

```


```{r, eval=FALSE}
## Another map
library("ggspatial")

maps.sites <- sites %>% 
  ggplot() +
  geom_sf(aes(color = Type))

maps.sites +
  annotation_scale(location = "bl", width_hint = 0.5) +
    annotation_north_arrow(location = "bl", which_north = "true", 
        pad_x = unit(0.1875, "in"), pad_y = unit(0.85, "in"),
        style = north_arrow_fancy_orienteering)

```


## Willow Height


```{r}
ht <- read_csv("output/processed_data/ht_combined20200107_1618.csv")

```


## Willow Production and Utilization


## Hydrologic Measurements - Manual

```{r}
wt.m <- read_csv("output/processed_data/water_table/manual_WTD_site_mean_20191212.csv")
```

```{r}
wt.m %>% 
  visdat::vis_dat()

```


## WT
```{r}
wt.m %>%
  filter(rwte_mean_late_summer < 0) %>% 
  filter(plot_type != "obs") %>% 
  ggplot(aes(yr, rwte_mean_late_summer)) +
  geom_line(aes(color = plot_type)) +
  geom_point(aes(color = plot_type)) +
  ylim(-150, 10) +
  facet_wrap(~site) +
  theme_minimal()

```


## Hydrologic Measurements - Dataloggers


## Soil Moisture Sensors

-EM50 dataloggers have a particular output format. Need to cleave off several of the top rows.
-Decagon soil VWC and temp


## Camera Trap Data

## Survey Data


## Air Temperature and Precipitation


## Long-term Climate Data