---
title: "Willow Height - Data Processing Report"
output:
  html_document: 
    fig_caption: yes
    fig_height: 8
    theme: journal
    toc: yes
    toc_depth: 2
    toc_float: no
  word_document: default
---

```{r setup, include=FALSE}
# knitr::opts_chunk$set(echo = TRUE)
library(knitr)
opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.align = 'center')
# opts_knit$set(root.dir=normalizePath('../')) # this is required if Rmd is nested below the project directory

# opts_chunk$set(fig.path = "../output/figures/") # corrected path and added dev. Needed to specify a subdirectory for figs

```

**Updated:** `r format(Sys.time(), '%d %B, %Y, %H:%m')`



This document summarizes data combined, wrangled, and exported in a separate Rmd file.


```{r,echo=FALSE}
suppressPackageStartupMessages(library(tidyverse))
library(fs)
suppressPackageStartupMessages(library(sf))
# library(raster)
suppressPackageStartupMessages(library(janitor))
suppressPackageStartupMessages(library(readxl))
# library(glue)
suppressPackageStartupMessages(library(mapview))
# library(ggmap)
# library(ggrepel)
suppressPackageStartupMessages(library(viridis))
library(ggExtra)
library(ggrepel)
library(DT)
# library(kableExtra)
# suppressPackageStartupMessages(library(compare))
suppressPackageStartupMessages(library(skimr)) ## some useful functions
suppressPackageStartupMessages(library(dataMaid))
library(lubridate)
library(gt)
```



```{r, functs_plot}
# Functions

## plotting
ggTile_yr_season_site2 <- function(df){
  df %>%
  group_by(yr, season, site2) %>%
  summarise(n= n()) %>%
  ggplot(aes(yr, site2)) +
  geom_tile(aes(fill = n), color = 'white') +
  viridis::scale_fill_viridis(option = "D") +
  facet_wrap(~season) +
  theme(axis.text.x = element_text(angle = 55, hjust = 1, size = 7)) +
  theme(axis.text.y = element_text(size = 8)) +
  facet_wrap(~season)}

ggTile_yr_season_site <- function(df){
  df %>%
  group_by(yr, season, site) %>%
  summarise(n= n()) %>%
  ggplot(aes(yr, site)) +
  geom_tile(aes(fill = n), color = 'white') +
  viridis::scale_fill_viridis(option = "D") +
  theme(axis.text.x = element_text(angle = 55, hjust = 1, size = 7)) +
  theme(axis.text.y = element_text(size = 8)) +
  facet_wrap(~season)} 

##
ggTile_yr_season_plot <- function(df){
  df %>%
  group_by(yr, plot, season, site) %>%
  tally() %>%
    ungroup() %>% 
  ggplot(aes(plot, site)) +
  geom_tile(aes(fill = n), color = 'white') +
  viridis::scale_fill_viridis(option = "B") +
  facet_wrap(~season) +
  theme(axis.text.x = element_text(angle = 55, hjust = 1, size = 9)) +
  theme(axis.text.y = element_text(size = 9)) +
  facet_wrap(yr~season)
  } 
#

ggTile_yr_season_plot_alt1 <- function(df){
  df %>%
  group_by(yr, plot, season, site) %>%
  summarise(n= n()) %>%
  ggplot(aes(yr, site)) +
  geom_tile(aes(fill = n), color = 'white') +
  viridis::scale_fill_viridis(option = "A") +
  # facet_wrap(~season) +
  facet_grid(season~plot) +
  theme(axis.text.x = element_text(angle = 55, hjust = 1, size = 9)) +
  theme(axis.text.y = element_text(size = 9))
} 

### ggpoint
ggPoint_yr_site_season_plot <- function(df){
  df %>%
  ggplot(aes(yr, pl_ht_cm)) +
    geom_point(aes(color = plot, shape = plot)) +
    facet_wrap(~site, ncol = 4) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 9)) +
  theme(axis.text.y = element_text(size = 8)) +
    viridis::scale_color_viridis(discrete = TRUE, option = "C") +
    labs(y = "Height (cm)")
} 


### gt table summary
tbqc_exp <- function(df){
  df %>%
    filter(plot != "obs") %>% 
    group_by(site,yr, season, plot) %>% 
    summarise(n = count(), ht_mean = mean(pl_ht_cm, na.rm = TRUE), ht_sd = sd(pl_ht_cm, na.rm = TRUE)) %>% 
    gt()
}

```



```{r}
## Read in data
# ht.combined <- read_csv("./output/processed_data/ht_combined20191221_2053.csv")

ht.combined <- read_csv("./output/processed_data/ht_combined20200107_1618.csv")

ht.combined <- ht.combined %>% 
  mutate(yr = as.integer(yr))

# ht.combined %>% View()

```

# Observation tallies

## Site tally - all sites/yrs

```{r}
ht.combined %>% 
  ggTile_yr_season_site()
```

## Site tally - exp sites/yrs
```{r, fig.height=3.5}
ht.combined %>%
  filter(plot != "obs") %>% 
  ggTile_yr_season_site() 

```


## Site tally - obs sites/yrs
```{r, fig.height=4.5}
ht.combined %>%
  filter(plot == "obs") %>% 
  ggTile_yr_season_site() 

```

```{r}
# ht.combined %>%
#   filter(plot == "obs") %>%   
#   ggPoint_yr_site_season_plot
```

## Site2 tally - exp sites/yrs
```{r, fig.height=4.5}
ht.combined %>% 
  filter(plot != "obs") %>%
  ggTile_yr_season_site2()
```

## Site2 tally - obs sites/yrs
```{r, fig.height=6.75}
ht.combined %>%
  filter(plot == "obs") %>%
  ggTile_yr_season_site2()
```

# Height plots and summaries
```{r}
## summarize by site2 and year using skimr
summary.by.site2 <- ht.combined %>% 
  filter(!is.na(site)) %>% 
  mutate(year = as.factor(yr)) %>% 
  group_by(site, season, plot, site2,yr) %>%
  nest() %>% 
  mutate(skim2 = map(data,skim)) %>% 
  unnest(skim2) %>% 
  ungroup()
```

## Median heights - exp plots, site, plot, yr

>Suspect value for spring eb2 in 2015. Filtered out.

```{r, eval=TRUE}
## line plot
pltly3 <- summary.by.site2 %>%
  mutate(yr = as.integer(yr)) %>% 
  filter(skim_variable == 'pl_ht_cm') %>%
  filter(plot != "obs") %>% 
  filter(!(site2 == "elk-dx" & yr == 2019)) %>% 
  filter(!(site2 == "eb2-cc" & numeric.p50 >200)) %>% 
  ggplot(aes(yr)) +
  geom_line(aes(y=numeric.p50, color = plot), alpha = .5) +
  geom_point(aes(y=numeric.p50, color = plot)) +
  geom_pointrange(aes(y = numeric.p50, ymin = numeric.p25, ymax = numeric.p75, color = plot), alpha = 0.51) +
  # facet_wrap(~site) +
  facet_grid(site~season) +
  theme_minimal() +
  labs(y="Median willow height (cm)", x = "Year", caption = "Willow height at experimental sites")
# ggsave("./output/median_willow_height_20191205_0924.png", width = 7, height = 5.5)

pltly3 %>% plotly::ggplotly()

```

## Median heights - exp plots, site2, yr

>interactive plot. hover over point for more info

```{r, fig.height=8}
## line plot
ggpl1 <- summary.by.site2 %>%
  mutate(yr = as.integer(yr)) %>% 
  filter(skim_variable == 'pl_ht_cm') %>%
  filter(!(site2 == "eb2-cc" & numeric.p50 >200)) %>% 
  filter(plot != "obs") %>% 
  ggplot(aes(yr)) +
  # geom_line(aes(y=numeric.p50, color = season, lty = season)) +
  geom_point(aes(y=numeric.p50, color = season, shape = season)) +
  # geom_pointrange(aes(y = numeric.p50, ymin = numeric.p25, ymax = numeric.p75, color = plot), alpha = 0.31) +
  facet_wrap(~site2, ncol = 4) +
  # facet_grid(site~season) +
  theme_minimal() +
  labs(y="Median willow height (cm)", x = "Year", caption = "Willow height at experimental sites")
# ggsave("./output/median_willow_height_20191205_0924.png", width = 7, height = 5.5)
# ggpl1

plotly::ggplotly(ggpl1)

```

> suspect point: wb-cx 2018 spring

## Median heights - exp plots by site2

>interactive plot. hover over point for more info

```{r}
##
pltly2 <- summary.by.site2 %>%
  select(yr, plot, site2, skim_variable, contains("numeric")) %>% 
  filter(!(site2 == "elk-dx" & yr == 2019)) %>% 
  filter(!(site2 == "eb2-cc" & numeric.p50 >200)) %>%  
  filter(skim_variable == "pl_ht_cm") %>% 
  filter(plot != "obs") %>%
  mutate(yr = as.integer(yr)) %>% 
  ggplot(aes(yr, site2)) +
  geom_tile(aes(fill = numeric.mean), color = 'gray70') +
  scale_fill_viridis() +
  theme_light() +
  labs(x = "Year", y = "", subtitle = "Mean willow height", fill = "Height (cm)", caption = getwd())

pltly2 %>% plotly::ggplotly()

# ggsave("./output/median_willow_height_HM_20191205_0924.png", width = 7, height = 7)

```

